\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}
\thispagestyle{plain}
The idea behind using machine learning was born from my motivation to dive deep into this topic. The idea to use machine learning in relation with glacier modeling however, developed by following a course from Alex Jarosch who had previously participated in the development of the model from \citet{Clarke2009}. A special thanks then goes to him for the inspiration about this work.

The development of this thesis however was only possible thanks to Fabien Maussion, who already had in his mind the idea of using statistical models to estimate glaciers ice thickness from observations. A huge thanks than goes to him, who allowed me to work on the topic i wanted to explore, and gave this work its research focus. He also helped me navigating the hassles of having to deal with geographical data and the tools to do so (OGGM), and guided me through the whole process needed to complete this work. 

On this regard a big thanks goes to the whole OGGM community who made this thesis possible with the creation of the OGGM model.

Thanks also to Reto Stauffer who helped with focusing the statistical part in the right direction.
\newline

Thanks to all the friends i met during these years of studies at the department of atmospheric sciences and in particular to: Antonio, Edoardo, Mattia and Riccardo.

Thanks to Kata and Luigi who encouraged me to start this master course during a climbing day!
\newline

I would then like to thank my family and in particular my mom who helped me correcting the grammatical part of the thesis. 

A final huge thanks goes to my wife Cristina who supported me through the whole process of this master course and made studying for me these years possible. She is the one without whom writing this thesis wouldn't have been possible. It's been wonderful writing it with you and Isabella by my side! Having you made this work a type 2 experience and made the quarantine circumstances of the covid-19 a very happy time for me.

%Now it is time to thank all people who have contributed to your work and who
%have supported you during your study. Do not forget to mention all relevant data
%providers and funding agencies (also provide the grant numbers).
